---
title: "R Notebook"
output: html_notebook
---

-   *Run section* by pressing *Ctrl+Enter*.
-   *Run line* by pressing *Shift+Enter*.
-   *Insert Chunk* by pressing *Ctrl+i*.
-   *Pipe operator* press *Ctrl+p* to filter or select conditions `%>%`
-   *Copy text* press *Ctrl+c*
-   *open file* press *Ctrl+o*
-   *save file* press *Ctrl+s*
-   *paste all* press *Ctrl+a*

# Functions overview

-   `c` = ...
-   `list` = ...
-   `matrix` = ...
-   `class` = ...
-   `sum` = ...
-   `mean` = ...
-   `colnames` = ...
-   \`rownames\`\` = ...
-   \`colSums\`\` = ...
-   \`rowSums\`\` = ...
-   `cbind` = ...
-   \`rbind\`\` = ...
-   `factors` = ...
-   `levels` = ...
-   `summary` = ...
-   `is.numeric` = ...
-   `identical` = ...

# Start

```{r}

```

# Basics

```{r}
# Addition
5 + 5 

# Subtraction
5 - 5 

# Multiplication
3 * 5

# Division
(5 + 5) / 2 

# Exponentiation
2^5

# Modulo
28 %% 6
```

# Variables

Assign the value to a variable

```{r}
# Assign
my_numeric <- 42
my_character <- "universe"
my_logical <- FALSE
```

Print out the value of the variable x

```{r}
# Print
my_numeric
my_character
my_logical
```

Check the class/type of the variable

```{r}
# Check class
class(my_numeric)
class(my_character)
class(my_logical)
```

# Vectors

## Construct a vector

Filling a new vector

```{r}
# Vectors
numeric_vector <- c(1, 10, 49)
character_vector <- c("a", "b", "c")
boolean_vector <- c(TRUE, FALSE, TRUE)

# Print
numeric_vector
character_vector
boolean_vector
```

Give the vector header names

```{r}
# Vector names
poker_vector <- c(140, -50, 20, -120, 240)
names(poker_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

# Print
poker_vector
```

## Manipulating a Vector

Summation with vectors

```{r}
# Vector sum
A_vector <- c(1, 2, 3)
B_vector <- c(4, 5, 6)
total_vector <- A_vector + B_vector
total_sum <- sum(total_vector)

# Print
total_vector
total_sum
```

Slicing a vector

```{r}
# Slicing
poker_wednesday <- poker_vector[3]
poker_midweek1 <- poker_vector[c(2,3,4)]
poker_midweek2 <- poker_vector[2:4]
poker_start <- poker_vector[c("Monday","Tuesday","Wednesday")]

# Print
poker_wednesday
poker_midweek1
poker_midweek2
poker_start
```

Selection by condition

```{r}
# Selection
selection_vector <- poker_vector > 0
poker_winning_days <- poker_vector[selection_vector]

# Print
selection_vector
poker_winning_days
```

## Calculations

Calculating the mean

```{r}
# Stats
mean(poker_start)
```

# Matrices

## Construct a Matrix

Template for filing a matrix

```{r}
# Construct a matrix with 3 rows that contain the numbers 1 up to 9
matrix(1:9, byrow = TRUE, nrow = 3)
```

Define column values

```{r}
# Construct star_wars_matrix
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)
box_office <- c(new_hope, empire_strikes, return_jedi)
```

Fill the matrix

```{r}
# Option 1: Adding column and row names
star_wars_matrix_1 <- matrix(box_office, byrow = TRUE, nrow = 3)

region <- c("US", "non-US")
titles <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")

colnames(star_wars_matrix_1) <- region
rownames(star_wars_matrix_1) <- titles

star_wars_matrix_1
```

```{r}
# Option 2: Adding column and row names
star_wars_matrix_2 <- matrix(box_office
                           , byrow = TRUE
                           , nrow = 3
                           , dimnames = list(titles, region))

star_wars_matrix_2
```

## Manipulating the matrix

Summation

```{r}
# Stats
worldwide_vector <- rowSums(star_wars_matrix_1)
worldwide_vector
```

```{r}
# Stats
total_revenue_vector <- colSums(star_wars_matrix_1)
total_revenue_vector
```

Binding of columns and rows

```{r}
# Bind COLUMNS in a single matrix
all_wars_cmatrix <- cbind(star_wars_matrix_1, worldwide_vector)
all_wars_cmatrix
```

```{r}
# Bind ROWS in a single matrix
all_wars_rmatrix <- rbind(star_wars_matrix_1, star_wars_matrix_2)
all_wars_rmatrix
```

Slicing and calculating the mean

```{r}
# Slicing
non_us_all <- all_wars_rmatrix[,2] # Second column
non_us_all
```

```{r}
non_us_some <- all_wars_rmatrix[1:2, 2] # First two rows, of the second column
non_us_some
```

```{r}
# Stats
mean(non_us_some)
```

## Calculations

Matrix multiplication is usually done with: %\*% (standard).

Multiplication of every element by each relating element is done with: \*

```{r}
# Mathematical: Multiplication
twice_non_us_some <- non_us_some *2 
```

Similarly is for applying a division

```{r}
# Mathematical: Division (as example)
halfe_non_us_some <- non_us_some / 2 
```

# Factors

## Construct a Factor

A Factor defines the unique values in a list.

```{r}
# Convert sex_vector to a factor
sex_vector <- c("Male", "Female", "Female", "Male", "Male")
factor_sex_vector <- factor(sex_vector)
factor_sex_vector
```

## Calculations

Example with temperatures

```{r}
# Temperature (indicate vector is of ORDINAL type, levels give you the values in the correct order)
temperature_vector <- c("High", "Low", "High","Low", "Medium")
factor_temperature_vector <- factor(temperature_vector, order = TRUE, levels = c("Low", "Medium", "High"))

# Print
factor_temperature_vector

# Calculation
temp_day1 = factor_temperature_vector[1] # High
temp_day4 = factor_temperature_vector[4] # Low
temp_day1 > temp_day4 # Was day 1 warmer than day 4? -> TRUE
```

The names of the factors can be changed

```{r}
# Code to build factor_survey_vector
survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)
factor_survey_vector
```

Now, apply the change of the 'F' and 'M' levels of above code

```{r}
# Change variable names of the whole vector
# F becomes Female, M becomes 
levels(factor_survey_vector) <- c("Female", "Male")

# Male
factor_survey_vector
```

Generate a summary of a vector. This summary provides the length, class and mode of the vector.

```{r}
# Summrize the vector
summary(survey_vector)
```

Generate a summary of a factor. This summary provides a number of observations for each factor

```{r}
# Summrize the factors
summary(factor_survey_vector)
```

# List of vectors and matrices

## Construct a list

Separated vectors and matrices defined:

```{r}
# Define vectors and matrices
my_vector <- 1:10 
my_matrix <- matrix(1:9, ncol = 3)
my_df <- mtcars[1:10,]
```

Constructing a list step by step:

```{r}
# Make list of vector and matix
my_list <- list(my_vector, my_matrix, my_df) # create list
names(my_list) <- c('vec','mat','my_df') # give names to lists in my_list

my_list
```

All notations at once:

```{r}
# Shortcut
my_list2 <- list(vec = my_vector, mat = my_matrix, my_df = my_df) # all at once
my_list2
```

## Manipulating a list

Selecting a single object in the list can be done by multiple options

```{r}
# Selecting only my_vector

# Option 1
my_list[1]

# Option 2
my_list['vec']

# Option 3 (shortcut)
my_list$vec

# Option 3 (elaborated filter)
my_list$vec[2]
```

# Boolean

TRUE or FALSE?

```{r}
# Basics
TRUE == TRUE
'hello' != 'goodbye'
2 != 22
70 > 7
70 >= 7
'hello' > 'goodbye' # Alphabetical order
TRUE > FALSE # 1 > 0
```

Comparisions

```{r}
# vectors
link <- c(16,1,9,6,2,17,14)
combi <- c(7,4,5,14,21,28,3)

# Calculations
link > 7
link <= combi # value comparison one by one
identical(link, combi)
```

```{r}
combi <- c(7,4,5,14,21,28,3)

# Range calculations
combi > 5 & combi < 15
combi > 15 | combi < 5
```

Relations:

```{r}
# Relations 'AND'
c(TRUE, TRUE, FALSE) & c(TRUE, FALSE, FALSE) # elementwise
c(TRUE, TRUE, FALSE) && c(TRUE, FALSE, FALSE) # all
```

```{r}
# Relations 'OR'
c(TRUE, TRUE, FALSE) | c(TRUE, FALSE, FALSE) # elementwise
c(TRUE, TRUE, FALSE) || c(TRUE, FALSE, FALSE) # all
```

Is correct type?

```{r}
# Numeric check
!TRUE
is.numeric(5) # TRUE
!is.numeric('hello') # TRUE
```
