---
title: "R Notebook"
output: html_notebook
---

-   *Run section* by pressing *Ctrl+Enter*.
-   *Run line* by pressing *Shift+Enter*.
-   *Insert Chunk* by pressing *Ctrl+i*.
-   *Pipe operator* press *Ctrl+p* to filter or select conditions `%>%`
-   *Copy text* press *Ctrl+c*
-   *open file* press *Ctrl+o*
-   *save file* press *Ctrl+s*
-   *paste all* press *Ctrl+a*

# Functions overview

-   `head` = ...
-   `str` = ...
-   `subset` = ...
-   `order` = ...
-   `select` = ...
-   `filter` = ...
-   `arrange` = Sort columns
-   `mutate` = Add new feature
-   `transmutate` = ...
-   `count` = ...
-   `group_by` = ...
-   `ungroup` = ...
-   `summarize` = mean, sum, median, min, max
-   `rename` = ...
-   `glimpse` = ...
-   `slice_max` = ...
-   `slice_min` = ...
-   `top_n` = ...
-   `lag` = ...

```{r}
# Install packages
install.packages("gapminder")

# Load packages
library('gapminder')
library('dplyr')
```

```{r}
# Get explaination of a function
glimpse(dataset)
?select_helpers
```

# Dataset

## Make a dataset

```{r}
# Give  headers a name
name <- c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", "Neptune")

# Data to store
type <- c("Terrestrial planet", "Terrestrial planet", "Terrestrial planet", 
          "Terrestrial planet", "Gas giant", "Gas giant", "Gas giant", "Gas giant")
diameter <- c(0.382, 0.949, 1, 0.532, 1.209, 9.449, 4.007, 3.883)
rotation <- c(58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67)
rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)

# Create a data frame from the vectors
planets_df <- data.frame(name, type, diameter, rotation, rings)
mars <- planets_df[4,]
```

## Basic information

```{r}
# Make general name
dataset <- planets_df
```

```{r}
# head dataset
head(dataset)

```

```{r}
# structure dataset
str(dataset)
```

And other general notations to sort and to order a list

```{r}
# Example 1 - numeric
a <- c(100, 10, 1000)
order(a)

# Example 2 - strings
az <- c('a', 'F', 'f', 'A', 'z')
order(az)
```

```{r}
# Example 3 - order based on a column
positions <- order(planets_df$diameter)
planets_df[positions,]
```

## Select columns

Selecting a column can be done by tree different notations

```{r}

# Option 1
diameter <- planets_df[,3]

# Option 2
diameter <- planets_df[,"diameter"]

# Option 3: short-cut
planets_df$diameter 
```

## Filter by conditions

Selecting columns with condition can also be applied in different ways

```{r}
# Select columns with False-True list
rings_vector <- planets_df$rings

# Show 'name' with True rings
planets_df[rings_vector, "name"]

# Show all columns with True rings
planets_df[rings_vector, TRUE]
```

```{r}
# Setting a condition
subset(planets_df, subset = (diameter < 1))
```

# Pipe operations

Below, different pipe operations are presented as general expression.

## select

```{r}
dataset %>%
  select(column1, column2, column4:column7, new_name = old_name)

dataset %>%
  select(contains('word'), starts_with('letter'), ends_with('letter'), matches('letter'), last_col() )

dataset %>%
  select(-column_to_remove)
```

## filter

```{r}
dataset %>%
  filter(column1 == value1, column2 = value2)

dataset %>%
  filter(column1 == value1) %>%
  arrange(desc(value2))

dataset %>%
  filter(value1 %in% c('value1', 'value2'))
```

## group_by

```{r}
dataset %>%
  group_by(column1_to_group, column2_to_group) %>%
  summarize(column_name1 = mean(column1), column_name2 = sum(column2))
  ungroup() # Make it able to do other aggregations on the same dataset. 

dataset %>%
  group_by(column1_to_group, column2_to_group) %>%
  slice_max(column1, n = int)
  slice_min(column2, n = int)
  
dataset %>%
  group_by(column1_to_group) %>%
  mutate(new_col = sum(column2))
```

## Others

```{r}
dataset %>%
  rename(new_name = old_name)

dataset %>%
  arrange(column1, desc(column2)) # default is ascending

dataset %>%
  mutate(column1 = value1/10) # overwrite column

dataset %>%
  mutate(column_name = value1/10) # add column

dataset %>%
  transmute(column1, column2 = column2/10) # select + mutate in once

variable_name <- dataset %>%
  filter(column1 == value1)

dataset %>%
  summarize(column_name1 = mean(column1), column_name2 = sum(column2))

dataset %>%
  count(column1, wt = column2, sort = TRUE)
```

## Intermezzo: lag

```{r}
v <- c(1,3,6,14)
sub_the_previous <- lag(v)
diff_consecative <- v - lag(v)
```
