---
title: "R Notebook"
output: html_notebook
---

-   *Run section* by pressing *Ctrl+Enter*.
-   *Run line* by pressing *Shift+Enter*.
-   *Insert Chunk* by pressing *Ctrl+i*.
-   *Pipe operator* press *Ctrl+p* to filter or select conditions `%>%`
-   *Copy text* press *Ctrl+c*
-   *open file* press *Ctrl+o*
-   *save file* press *Ctrl+s*
-   *paste all* press *Ctrl+a*

# Functions overview

-   `function` = ...

# Start

```{r}
# Install packages
install.packages("ggplot2")

# Load packages
library('ggplot2')
```

# Visualisations

## Scatterplot

Single plot

```{r}
# Basics
ggplot(data, aes(x = column_x, y = column_y)) + # aes = aesthetic
  geom_point() # adding a layer/ type_of_graph, point --> scatter
```

```{r}
# Designing 
ggplot(data, aes(x = column_x, y = column_y, color = column_col, size = column_size)) + 
  geom_point() +
  scale_x_log10() +                         # x-axis on a log scale
  scale_y_log10() +                         # y-axis on a log scale
  expand_limits(y=0) +                      # set limits
  ggtitle("Title") +
  labs(x = "label_x", y = "label_y")
```

Multiple plots by subplots (=faceting)

```{r}
# Basics
ggplot(data, aes(x = column_x, y = column_y, color = column_col, size = column_size)) + 
  geom_point() +
  facet_wrap(~ column_to_facet)
```

## Lineplot

single plot

```{r}
# Basics
ggplot(data, aes(x = column_x, y = column_y, color = column_col, size = column_size)) + 
  geom_line()
```

## Barplot

single plot

```{r}
# Basics
ggplot(data, aes(x = column_x, y = column_y, color = column_col, size = column_size)) + 
  geom_col()
```

## Histogram

single plot

```{r}
# Basics
ggplot(data, aes(x = column_x,  color = column_col, size = column_size)) + 
  geom_histogram(binwidth = int, bins = int)
```

## Boxplot

single plot

```{r}
# Basics
ggplot(data, aes(x = column_x, y = column_y, color = column_col, size = column_size)) + 
  geom_boxplot()
```
